Performance of two different ontological and data access methods

### 2014-07-02 cold2014/magnus.tex
mention Lenths method
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@715 43ef1190-90fd-4dd7-a769-d0121f02b80e

We present two different approaches. One is promoted by Renault, see
\cite{SemWebAppRes} and \cite{ren1}. They have developed the
Configuration Ontology (CO) \cite{confOnt}, and developed an API to allow
users to traverse the RDF graph to find a smaller result set of cars.
The ontology focuses on building an RDF graph to represent all valid
configurations for a product model with component constraints.  It is
also made so that it is not needed to do any reasoning to get a valid
result set.  They call it a traversal of \emph{Configurations} which
will end up with a \emph{Partially Defined Products}
(PDP)\footnote{Partially Defined Product is a way of defining a
  product without using all of the features. For instance a car which
  may have many features, but we describe it as a car with mp3 and
  sunroof.}.  Their method insists there is a starting point, which is
a car model. From there, there is a \emph{lexicon} which can be used
to choose different specifications which may lead to a valid
configuration. There are also stated if some specifications are
impossible in the current configuration.  These are represented by
either the \emph{co:possible} or \emph{co:impossible}
property. Furthermore, string comparison must be used to check if a
certain specification is representing for instance fuel type.

@@

In the experiments the definition for VSO/COO is \emph{-vso} and for
CO is \emph{-co}.  These two definitions stand for which ontology
are tested in a particular run. In practice this means that when we
send in the keyword \emph{-vso} we only run the specification search
against the car models represented by VSO/COO, in our case the data
from Daimler. This is referred to as the ontology factor in each
experiment and is a factor in all the experiments which are presented
in this thesis. The ontology factor represents the difference in how
the data are represented and how the ontologies are structures, and
with this factor we want to determine how this affects the overall
performance.  We have programmed the application so that the querying
of the ontologies are done as similar as possible on both levels. This
was done to eliminate any interference from the application. There are
some differences in the programming due that the ontologies were
different and they will be discussed later on.  In Chapter
\ref{prototype}, the application and its differences are explained in
depth.

Before starting on the experiment we had to find a way to do the
actual experiment. We wanted to first create the planning matrix and
then execute each run as a HTTP post. To do this we needed to make a
small script to calculate the $2^k$ runs and perform the HTTP
posts. We chose to use the programming language \emph{python} to do
this. That was because python was a familiar language and it also had
a package for DoE\footnote{http://pythonhosted.org/pyDOE/}. With the
DoE package we just declared the amount of factors and the matrix was
created. At that point the matrix looked more like Table
\ref{designs}, which meant that we had to substitute the + and - with
the levels of each factor.  Further in the script we used the package
\emph{request}\footnote{http://requests.readthedocs.org/en/latest/}
to do the HTTP posts in python. Below one can see the line of code
executing the request.

@@

The \emph{payload} is a dictionary of all the values to post to the
application.  After the request we saved the response which contained
the response time of a run. After each request we saved the run plus
the response time and wrote it to file on CSV format. To evaluate the
result we used the programming language R with the packages
\emph{DoE.base}~\cite{DoEBase}, \emph{FrF2}~\cite{FrF2} and
\emph{BsMD}~\cite{BsMD}. R allowed us to parse the CSV files created
from the experiments and output them in two types of graphs.  A
similarity with all the experiments were that they all contained at
least the factors ontology, transmission and fuel type.

To evaluate the graphs we have used a method in the \emph{FrF2} R
package called \emph{DanielPlot}. This method assumes that the
estimated effects are normally distributed with means equal to the
effects. The means of all estimated effects are zero. Resulting in a
plot where the estimated effect would end up on a straight line. This
plot is then testing whether all the estimated effects have the same
distribution. All deviations of the straight line indicates a
significant factor. In our results we have used the absolute value of
the effects, also called a half-normal plot. The advantages of using
the half-normal plot is that all large estimated effects will end up
in the top right corner. This means that the highest performance
significance will end up in
the top right corner.~\cite{Plotting} \\
Here is a quick explanation of the axes shown in each graph.  The X
axis will show the absolute effect of each factor, shown in time. The
values on the X axis is represented in seconds which means that the a
plot will have a significance measured in seconds.  The Y axis shows
the half-normal scores. There are two way of representing the graph
and that is with either half-normal plots or normal plots. With
half-normal the effects on the X axis are shown with the absolute
effect. This means that all factors are placed with a positive
half-normal score. This is to avoid visual misleading scores because
with the normal scores, deviations from the line on negative scores
might confuse the reader who are evaluating the charts.  The scores
them selves are a measure for which factor that is the most
significant.  We have also set the \emph{alpha} to $0.05$ which
means that we will tolerate a false percentage of $5\%$.

@@

Like the rest of the experiments we used DanielPlot to evaluate and a
package in R to visualize the result. The Figure \ref{3factorGraph}
shows the results.  Here we see no significant factors. Still there
are small effects from several factors. In Table \ref{3factorEffect}
we can see the time effects from largest to smallest.  We see that the
ontology factor has the largest effect on approximately 2.2 seconds
which means that one of the ontologies overall is 2.2 seconds
faster. By reading the results from the linear model it is the factor
level \emph{-vso} which has an overall faster response time.

@@

In the graph \ref{10factorGraph} we can see the results from running
all the 1024 runs.  We can see that there are a lot of significant
effects present here. There are more than any of the previous
experiments, but it is also more factors and factor interactions.
Unfortunately there are so many significant factors that there are
some of them that are illegible, but most of them have a low impact on
the significance graph. Like the six factor experiment we also see
here that there are three groups of significant factors. The first
group where most of them are placed range from around 2 seconds effect
to 11. Here there are a lot of factor interactions and why they are
there will be explained in the discussion.  The second group is very
like the second group we saw in the six factor experiment, the only
difference is that the total weight (WT) factor is substituted with
fuel consumption (FC). The last group is the same for all the
experiments, except the two factor experiment. Here we find the
ontology factor.  In Table \ref{10factorEffect} we can see the top
twenty absolute effects. This includes all the significant factors
from the last group and the second one, but also 5 from the first
group.  We can see that emission also here are present in a lot of the
entries, but so is ontology as well. The new thing here is the fuel
consumption has entered the most significant factors, and is by itself
higher than emission. The linear model again tells us that the option
\emph{-vso} has a significantly lower response time than
\emph{-co}

@@

If we take another look at Table \ref{10factorEffect} we see that
there are only three single factors except O that are in the top 15
significant factors.  After the 15 significant factor the absolute
effect drops significantly. The three factors are E, FC and T. E and T
are present because of the same reasons as in the six factor
experiment. We also see in the top 15 that there are only these three
factors present in all the factor interactions as well, which
indicates that T, FC and E affects each other. This is to be expected
because if a car has a high fuel consumption it also has a high
emission rate. There are some possible explanations for why there are
so many significant factor interactions in some of the
experiments. The order will affect this as mention above. Another
explanation is that we chose an average value to be an upper and lower
limit of a factor. We used these as factor levels to avoid an
unrealistic high number of levels. This could have opened for more new
specifications to check on one level of the factor.  As we see in the
Table \ref{10factorEffect}, \emph{E:WT} is higher than \emph{WT}
itself. This might indicate that one of the levels of E will open up
for more values to traverse in one of the levels of WT. This is a
complex matter which will depend on each product model as well as the
interaction between the specifications.

@@

The ontologies in this thesis are not made for tackling every possible
problem, but can perform well within their domain. The ontology
structure of CO in conjunction of the data representation, is made for
one particular scenario. That is to specify one specification after
another, not all at once. Here CO has several tools at disposal to
make it easier trimming down the result sets. The property
\emph{co:impossible} was to no use in our application due that our
application specified all specifications before searching. This
property lets the application remove possibilities if a user have
chosen some specifications. One would still has to use the lexicon as
reference, but could eliminate further possibilities. The property
\emph{co:impliedSpec} does something similar and gives the user a
list of specifications that are implied and do not need to be searched
for.  COO have some of the same functions and possibilities, but some
of these has to be reasoned over to be useful. One can produce the
same functionality by using \emph{coo:incompatibleWith} with
reasoning.

>>>>

We present two different approaches. One is promoted by Renault, see
\cite{SemWebAppRes} and \cite{ren1}. They have developed the
Configuration Ontology (CO) \cite{confOnt}, and developed an API to allow
users to traverse the RDF graph to find a smaller result set of cars.
The ontology focuses on building an RDF graph to represent all valid
configurations for a product model with component constraints.  It is
also made so that it is not needed to do any reasoning to get a valid
result set.  They call it a traversal of \emph{Configurations} which
will end up with a \emph{Partially Defined Products}
(PDP)\footnote{Partially Defined Product is a way of defining a
  product without using all of the features. For instance a car which
  may have many features, but we describe it as a car with mp3 and
  sunroof.}.  Their method insists there is a starting point, which is
a car model. From there, there is a \emph{lexicon} which can be used
to choose different specifications which may lead to a valid
configuration. There are also stated if some specifications are
impossible in the current configuration.  These are represented by
either the \textsf{co:possible} or \textsf{co:impossible}
property. Furthermore, string comparison must be used to check if a
certain specification is representing for instance fuel type.

@@

In the experiments the definition for VSO/COO is \textsf{-vso} and for
CO is \textsf{-co}.  These two definitions stand for which ontology
are tested in a particular run. In practice this means that when we
send in the keyword \textsf{-vso} we only run the specification search
against the car models represented by VSO/COO, in our case the data
from Daimler. This is referred to as the ontology factor in each
experiment and is a factor in all the experiments which are presented
in this thesis. The ontology factor represents the difference in how
the data are represented and how the ontologies are structures, and
with this factor we want to determine how this affects the overall
performance.  We have programmed the application so that the querying
of the ontologies are done as similar as possible on both levels. This
was done to eliminate any interference from the application. There are
some differences in the programming due that the ontologies were
different and they will be discussed later on.  In Chapter
\ref{prototype}, the application and its differences are explained in
depth.

Before starting on the experiment we had to find a way to do the
actual experiment. We wanted to first create the planning matrix and
then execute each run as a HTTP post. To do this we needed to make a
small script to calculate the $2^k$ runs and perform the HTTP
posts. We chose to use the programming language \emph{python} to do
this. That was because python was a familiar language and it also had
a package for DoE\footnote{http://pythonhosted.org/pyDOE/}. With the
DoE package we just declared the amount of factors and the matrix was
created. At that point the matrix looked more like Table
\ref{designs}, which meant that we had to substitute the + and - with
the levels of each factor.  Further in the script we used the package
\textsf{request}\footnote{http://requests.readthedocs.org/en/latest/}
to do the HTTP posts in python. Below one can see the line of code
executing the request.

@@

The \emph{payload} is a dictionary of all the values to post to the
application.  After the request we saved the response which contained
the response time of a run. After each request we saved the run plus
the response time and wrote it to file on CSV format. To evaluate the
result we used the programming language R with the packages
\textsf{DoE.base}~\cite{DoEBase}, \textsf{FrF2}~\cite{FrF2} and
\textsf{BsMD}~\cite{BsMD}. R allowed us to parse the CSV files created
from the experiments and output them in two types of graphs.  A
similarity with all the experiments were that they all contained at
least the factors ontology, transmission and fuel type.

To evaluate the graphs we have used a method in the \textsf{FrF2} R
package called \textsf{DanielPlot}. This method assumes that the
estimated effects are normally distributed with means equal to the
effects. The means of all estimated effects are zero. Resulting in a
plot where the estimated effect would end up on a straight line. This
plot is then testing whether all the estimated effects have the same
distribution. All deviations of the straight line indicates a
significant factor. In our results we have used the absolute value of
the effects, also called a half-normal plot. The advantages of using
the half-normal plot is that all large estimated effects will end up
in the top right corner. This means that the highest performance
significance will end up in
the top right corner.~\cite{Plotting} \\
Here is a quick explanation of the axes shown in each graph.  The X
axis will show the absolute effect of each factor, shown in time. The
values on the X axis is represented in seconds which means that the a
plot will have a significance measured in seconds.  The Y axis shows
the half-normal scores. There are two way of representing the graph
and that is with either half-normal plots or normal plots. With
half-normal the effects on the X axis are shown with the absolute
effect. This means that all factors are placed with a positive
half-normal score. This is to avoid visual misleading scores because
with the normal scores, deviations from the line on negative scores
might confuse the reader who are evaluating the charts.  The scores
them selves are a measure for which factor that is the most
significant.  We have also set  the significance level at $\alpha =
0.05$, where \textsf{DanielPlot} is using Lenth's method for the
analysis of unreplicated factorials. % TODO: Need a citation here.

@@

Like the rest of the experiments we used DanielPlot to evaluate and a
package in R to visualize the result. The Figure \ref{3factorGraph}
shows the results.  Here we see no significant factors. Still there
are small effects from several factors. In Table \ref{3factorEffect}
we can see the time effects from largest to smallest.  We see that the
ontology factor has the largest effect on approximately 2.2 seconds
which means that one of the ontologies overall is 2.2 seconds
faster. By reading the results from the linear model it is the factor
level \textsf{-vso} which has an overall faster response time.

@@

In the graph \ref{10factorGraph} we can see the results from running
all the 1024 runs.  We can see that there are a lot of significant
effects present here. There are more than any of the previous
experiments, but it is also more factors and factor interactions.
Unfortunately there are so many significant factors that there are
some of them that are illegible, but most of them have a low impact on
the significance graph. Like the six factor experiment we also see
here that there are three groups of significant factors. The first
group where most of them are placed range from around 2 seconds effect
to 11. Here there are a lot of factor interactions and why they are
there will be explained in the discussion.  The second group is very
like the second group we saw in the six factor experiment, the only
difference is that the total weight (WT) factor is substituted with
fuel consumption (FC). The last group is the same for all the
experiments, except the two factor experiment. Here we find the
ontology factor.  In Table \ref{10factorEffect} we can see the top
twenty absolute effects. This includes all the significant factors
from the last group and the second one, but also 5 from the first
group.  We can see that emission also here are present in a lot of the
entries, but so is ontology as well. The new thing here is the fuel
consumption has entered the most significant factors, and is by itself
higher than emission. The linear model again tells us that the option
\textsf{-vso} has a significantly lower response time than
\textsf{-co}

@@

If we take another look at Table \ref{10factorEffect} we see that
there are only three single factors except O that are in the top 15
significant factors.  After the 15 significant factor the absolute
effect drops significantly. The three factors are E, FC and T. E and T
are present because of the same reasons as in the six factor
experiment. We also see in the top 15 that there are only these three
factors present in all the factor interactions as well, which
indicates that T, FC and E affects each other. This is to be expected
because if a car has a high fuel consumption it also has a high
emission rate. There are some possible explanations for why there are
so many significant factor interactions in some of the
experiments. The order will affect this as mention above. Another
explanation is that we chose an average value to be an upper and lower
limit of a factor. We used these as factor levels to avoid an
unrealistic high number of levels. This could have opened for more new
specifications to check on one level of the factor.  As we see in the
Table \ref{10factorEffect}, \textsf{E:WT} is higher than \textsf{WT}
itself. This might indicate that one of the levels of E will open up
for more values to traverse in one of the levels of WT. This is a
complex matter which will depend on each product model as well as the
interaction between the specifications.

@@

The ontologies in this thesis are not made for tackling every possible
problem, but can perform well within their domain. The ontology
structure of CO in conjunction of the data representation, is made for
one particular scenario. That is to specify one specification after
another, not all at once. Here CO has several tools at disposal to
make it easier trimming down the result sets. The property
\textsf{co:impossible} was to no use in our application due that our
application specified all specifications before searching. This
property lets the application remove possibilities if a user have
chosen some specifications. One would still has to use the lexicon as
reference, but could eliminate further possibilities. The property
\textsf{co:impliedSpec} does something similar and gives the user a
list of specifications that are implied and do not need to be searched
for.  COO have some of the same functions and possibilities, but some
of these has to be reasoned over to be useful. One can produce the
same functionality by using \textsf{coo:incompatibleWith} with
reasoning.
